---
title: "GAM with Super Learner smoother"
author: "Jeremy Coyle"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
bibliography: references.bib
vignette: > 
  %\VignetteIndexEntry{gasl}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
library(gasl)
```

`gasl` is a package for fitting a GAM with SuperLearner smoothers for one or more components. Currently, it is structured for two smoothers, one for a "factor" that is a combination of covariates and a second for covariates that aren't part of the factor. This could be used to create an index variable such as a wealth index. This index variable can then be used in future studies, or as part of a data adaptive target parameter [@Hubbard:2016bg]. As compared to unsupervised apporaches such as PCA, this method is focused on generating an index variable that is predictive of the outcome of interest.

To demonstrate the package usage, we've included some simulated data:
```{r data}
data(gasl_test_sim)
str(gasl_test_sim$sim_data)
```
_todo: writeup DGD_

To fit a GASL model, simply specify the covariate matrix and outcome vector, and a character vector of covariates to include in the factor. Any other covariates will be fit by a separate smoother. By default, both the factor and covariates will be fit using a linear model. By using `factor_args` or `covariate_args` to specify a list of arguments to `origami::orgami_SuperLearner`, including SuperLearner library and fold structure, either term can be fit using a SuperLearner. 

```{r gasl_call}
#construct gasl arguments
X <- gasl_test_sim$sim_data[, c("A", "W1", "W2", "W3")]
X$AW1 <- X$A * X$W1
X <- as.matrix(X)
Y <- gasl_test_sim$sim_data$Y

factor_vars <- c("W2", "W3")
covariate_args <- list(SL.library = c("SL.glm", "SL.mean"), folds = make_folds(Y))

#fit gasl
gasl_fit <- gasl(X, Y, factor_vars, covariate_args = covariate_args)

#inspect results
gasl_fit

```

Once the GASL model is fit, predictions can be generated on new datasets.

```{r gasl_predict}
test_data <- gasl_test_sim$test_data
test_X <- test_data[, c("A", "W1", "W2", "W3")]
test_X$AW1 <- test_X$A * test_X$W1
```

```{r gasl_predict2}
test_data$gasl_pred <- predict(gasl_fit, newdata = test_X, type = "response")
```

```{r gasl_predict3}
gasl:::logit_risk(test_data$Y, test_data$null_pred)
gasl:::logit_risk(test_data$Y, test_data$gasl_pred)
gasl:::logit_risk(test_data$Y, test_data$true_pred)
gasl:::logit_risk(test_data$Y, test_data$Q0)
```
